---
title: "Lab 7: MATH 4753"
author: "Hunter DeVoe"
date: "February 8, 2023"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(pacman, tidyverse, knitr, devtools)

mychisim<-function(n1,sigma1,mean1,iter,ymax,x,y){    # adjust ymax to make graph fit
  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1

  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1

  ssq1=apply(data1.mat,2,var) # ssq1 is s squared

  w=(n1-1)*ssq1/sigma1^2      #chi-sq stat

  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
  main=substitute(paste("Sample size = ",n[1]," = ",n1," statistic = ",chi^2)),
  xlab=expression(paste(chi^2, "Statistic",sep=" ")), las=1)
  lines(density(w),col="Blue",lwd=3) # add a density plot
  curve(dchisq(x,n1-1),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  title=expression(chi^2==frac((n[1]-1)*s^2,sigma^2)) #mathematical annotation -see ?plotmath
  legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
  return(invisible(list(w=w,summary=summary(w),sd=sd(w),fun="Chi-sq"))) # some output to use if needed
}

myTsim<-function(n1,sigma1,mean1,iter,ymax,x,y){    # adjust ymax to make graph fit
  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
  
  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
  
  sd1=apply(data1.mat,2,sd) # sd
  ybar=apply(data1.mat,2,mean)  # mean
  
  w=(ybar-mean1)/(sd1/sqrt(n1))      #T stat
  
  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
  main=substitute(paste("Sample size = ",n[1]," = ",n1," statistic = ",T," iterations= ",iter)),
  xlab=expression(paste(T, "Statistic",sep=" ")), las=1)
  lines(density(w),col="Blue",lwd=3) # add a density plot
  curve(dt(x,n1-1),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  title=expression(T==frac((bar(y)-mu),s/sqrt(n1))) #mathematical annotation -see ?plotmath
  legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
  invisible(list(w=w,summary=summary(w),sd=sd(w),fun="T")) # some output to use if needed
 return(invisible(list(w=w,summary=summary(w),sd=sd(w),fun="T"))) # some output to use if needed
}

mychisim2<-function(n1,n2,sigma1,sigma2,mean1,mean2,iter,ymax,x,y){    # adjust ymax to make graph fit
  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
  y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
  data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
  ssq1=apply(data1.mat,2,var) # ssq1 is s squared
  ssq2=apply(data2.mat,2,var)
  spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
  w=(n1+n2-2)*spsq/(sigma1^2)#sigma1=sigma2,  Chi square stat
  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
  main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",chi^2)),
  xlab=expression(paste(chi^2, "Statistic",sep=" ")), las=1)
  lines(density(w),col="Blue",lwd=3) # add a density plot
  curve(dchisq(x,n1+n2-2),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  title=expression(chi^2==frac((n[1]+n[2]-2)*S[p]^2,sigma^2)) #mathematical annotation -see ?plotmath
  legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
  return (invisible(list(w=w,summary=summary(w),sd=sd(w),fun="Chi-sq"))) # some output to use if needed
}

myTsim2<-function(n1,n2,sigma1,sigma2,mean1,mean2,iter,ymax,x,y){
  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
  y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
  data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
  ssq1=apply(data1.mat,2,var) # ssq1 is s squared
  ybar1= apply(data1.mat,2,mean)
  ssq2=apply(data2.mat,2,var)
  ybar2=apply(data2.mat,2,mean)
  spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
  w=((ybar1-ybar2)-(mean1-mean2))/sqrt(spsq*(1/n1+1/n2))#sigma1=sigma2,  Chi square stat
  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
  main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",T)),
  xlab=paste(" T Statistic",sep=""), las=1)
  lines(density(w),col="Blue",lwd=3) # add a density plot
  curve(dt(x,n1+n2-2),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  title=expression(T==frac((bar(Y)[1]-bar(Y)[2])-(mu[1]-mu[2]),S[p]*sqrt(frac(1,n[1])+frac(1,n[2])))) #mathematical annotation -see ?plotmath
  legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
  return(invisible(list(w=w,summary=summary(w),sdw=sd(w),fun="T"))) # some output to use if needed
}

myFsim2<-function(n1,n2,sigma1,sigma2,mean1,mean2,iter,ymax,x,y){
  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
  y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
  data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
  ssq1=apply(data1.mat,2,var) # ssq1 is s squared
  ssq2=apply(data2.mat,2,var)
  #spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
  w=ssq1*sigma2^2/(ssq2*sigma1^2) #
  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
  main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",F)),
  xlab=paste("F Statistic",sep=""), las=1)
  lines(density(w),col="Blue",lwd=3) # add a density plot
  curve(df(x,n1-1,n2-1),xlim=c(0,6),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  title=expression(F==frac(s[1]^2,s[2]^2)*frac(sigma[2]^2,sigma[1]^2)) #mathematical annotation -see ?plotmath
  legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
  return(invisible(list(w=w,summary=summary(w),sd=sd(w),fun="F"))) # some output to use if needed
}
```

# Task 1

```{r}
getwd()
```

# Task 2

## Chi-Square Statistic

```{r}
mychisim(n1 = 10, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.12, x = 20, y = 0.08)

mychisim(n1 = 20, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.08, x = 30, y = 0.06)

mychisim(n1 = 100, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.04, x = 120, y = 0.03)

mychisim(n1 = 200, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.03, x = 225, y = 0.025)
```

## Make a Histogram of chisq$w

```{r}
chiwval <- mychisim(n1 = 10, sigma1 = 10, mean1 = 20, iter = 1500, ymax = 0.12, x = 20, y = 0.07)
hist(chiwval$w, col = rainbow(16), xlab = expression(paste(chi^2, "Value", sep = " "), las = 1))
```


# Task 3

## T Statistic

```{r}
myTsim(n1 = 10, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.5, x = 2, y = 0.3)

myTsim(n1 = 20, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.5, x = 2, y = 0.3)

myTsim(n1 = 100, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.5, x = 1.5, y = 0.4)

myTsim(n1 = 200, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.5, x = 1.5, y = 0.4)
```

## Make a Histogram of myTsim$w

```{r}
twval <- myTsim(n1 = 10, sigma1 = 10, mean1 = 20, iter = 1500, ymax = 0.4, x = 20, y = 0.07)
hist(twval$w, col = rainbow(16), xlab = expression(paste(T, "Value", sep = " "), las = 1))
```


# Task 4

## Two Sample Chi-Square Statistic

```{r}
mychisim2(n1 = 10, n2 = 10, sigma1 = 4, sigma2 = 4, mean1 = 5, mean2 = 10, iter = 1000, ymax = 0.10, x = 30, y = 0.08)

mychisim2(n1 = 20, n2 = 10, sigma1 = 10, sigma2 = 10, mean1 = 3, mean2 = 5, iter = 1000, ymax = 0.10, x = 40, y = 0.08)

mychisim2(n1 = 50, n2 = 50, sigma1 = 4, sigma2 = 4, mean1 = 5, mean2 = 10, iter = 10000, ymax = 0.04, x = 130, y = 0.03)

mychisim2(n1 = 80, n2 = 50, sigma1 = 10, sigma2 = 10, mean1 = 3, mean2 = 5, iter = 10000, ymax = 0.03, x = 160, y = 0.02)
```

## Make a Histogram of mychisim2$w

```{r}
chisim2wval <- mychisim2(n1 = 10, n2 = 14, sigma1 = 3, sigma2 = 3, mean1=  5, mean2 = 10, iter = 10000, ymax = 0.07, x = 40, y = 0.05)
hist(chisim2wval$w, col = rainbow(16), xlab = expression(paste(T, "Value", sep = " "), las = 1))
```


# Task 5

## Two Sample T Statistic

```{r}
myTsim2(n1 = 10, n2 = 10, sigma1 = 4, sigma2 = 4, mean1 = 5, mean2 = 10, iter = 1000, ymax = 0.5, x = -4, y = 0.4)

myTsim2(n1 = 20, n2 = 10, sigma1 = 10, sigma2 = 10, mean1 = 3, mean2 = 5, iter = 1000, ymax = 0.5, x = -4, y = 0.4)

myTsim2(n1 = 50, n2 = 50, sigma1 = 4, sigma2 = 4, mean1 = 5, mean2 = 10, iter = 10000, ymax = 0.5, x = -4, y = 0.4)

myTsim2(n1 = 80, n2 = 50, sigma1 = 10, sigma2 = 10, mean1 = 3, mean2 = 5, iter = 10000, ymax = 0.5, x = -4, y = 0.4)
```

## Make a Histogram of myTsim2$w

```{r}
myTsim2wval <- myTsim2(n1=10,n2=14,sigma1=3,sigma2=3,mean1=5,mean2=10,iter=10000,ymax=0.5,x=-5,y=0.4)
hist(myTsim2wval$w, col = rainbow(16), xlab = expression(paste(T, "Value", sep = " "), las = 1))
```


# Task 6

## Use the table to write down the statistic that the function will calculate  

$(\frac{S^2_1}{S^2_2}) (\frac{(\sigma^2_1)}{\sigma^2_2})$  

## What assumptions are made?  

For the Sampling Distribution: $F$ distribution with $v_1 = (n_1 - 1)$ numerator degrees of freedom and $v_2 = (n_2 - 1)$ denominator degrees of freedom.  

Additional Assumptions: None

## Two Sample F Statistic

```{r}
myFsim2(n1 = 50, n2 = 50, sigma1 = 4, sigma2 = 4, mean1 = 5, mean2 = 10, iter = 1000, ymax = 1.7, x = 2, y = 1.25)

myFsim2(n1 = 20, n2 = 80, sigma1 = 10, sigma2 = 10, mean1 = 3, mean2 = 5, iter = 1000, ymax = 1.7, x = 2, y = 1.25)

myFsim2(n1 = 50, n2 = 50, sigma1 = 4, sigma2 = 4, mean1 = 5, mean2 = 10, iter = 10000, ymax = 1.7, x = 2, y = 1.25)

myFsim2(n1 = 20, n2 = 80, sigma1 = 10, sigma2 = 10, mean1 = 3, mean2 = 5, iter = 10000, ymax = 1.7, x = 2, y = 1.25)
```

## Make a Histogram of myFsim2$w

```{r}
myFsim2wval <- myFsim2(n1=10,n2=14,sigma1=3,sigma2=2,mean1=5,mean2=10,iter=1000,ymax=0.9,x=3,y=0.6)
hist(myFsim2wval$w, col = rainbow(16), xlab = expression(paste(T, "Value", sep = " "), las = 1))
```

# Task 7

```{r}
devtools::install_local("C:/Users/Hunter/Desktop/Math4753Spring2023.zip")
```

```{r}
library(Math4753Spring2023)
```

```{r}
data("fire")
knitr::kable(head(fire))
```


```{r}
# CLEAN UP #################################################
# Clear environment
rm(list = ls()) 
# Clear packages
p_unload(all)  # Remove all add-ons
# Clear plots
dev.off()  # But only if there IS a plot
# Clear console
cat("\014")  # ctrl+L
# Clear mind :)
```